---
title: "Lecture 5"
author: "Ben Ridenhour"
date: "2022-09-08"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

- We picked up our discussion of the factor class again
- Using (statistical) formulae in various functions (e.g.. `lm()`)
    + See `? formula` for much of what I talked about
- We looked at differences in model output (and matrices) when changing the order of factor levels
- Appending new elements to factors
    + Level must exist OR 
    + `c(a,b)` works if both are factors
- Using the `sample()` and `set.seed()` functions
- Creating ordinal factors (with an example of ordinal logistic regression)
- The `tapply()` function
- The Date S3 class
- The `unclass()` and `class()` attributes functions

______________________________

# Quiz

1.  Create a vector `x` containing the numbers 4.2, pi, 2/3, and 3 + 2 i. Coerce the vector to be have the type `integer` and save it as `y`. Use a built-in R function to test whether the coercion worked. Explain the resulting resulting vector.

```{r}
x <- c(4.2, pi, 2/3, 3+2i)
y <- as.integer(x)
```

*The resulting vector happens because, when coerced to an integer, double values lose all decimal places and complex values lose their imaginary part.*

2.  By altering the attributes of `y`, create a 2 x 2 matrix. (In other words, **do not** use the `matrix()` command.) What are 2 other ways to alter the attributes to get the same 2 x 2 matrix?

```{r}
dim(y) <- c(2,2)
dim(y) <- NULL

attr(y, "dim") <- c(2,2)
y <- as.vector(y)

structure(y, dim = c(2,2))
```

3.  If `y` is no longer a vector, make it a vector again. Add the number 10 in the 20th element/position of `y`. Run the function `sum()` on `y`. Why does it fail? Look at the help for the sum function to figure out how to correct the issue. Can you figure out a way to ask R to count the number of **NA** elements for you?

```{r}

y[[20]] <- 10
sum(y)
sum(y, na.rm = T)

sum(is.na(y)) #counting NA values
```

4.  Create a factor `z` that has the values "a", "truck", and "100" repeated four times each. What attributes does `z` have? Which one of the attributes makes `z` an S3 object? Run the command `w <- unclass(z)`. Run `print(w)` and note the output. Run `class(w) <- "factor"`. Now `print(w)`. Explain why there is a difference.

```{r}
z <- factor(rep(c("a","truck","100"),4))
attributes(z)
w <- unclass(z)
w
class(w) <- "factor"
```

*The difference occurs because the generic function `print()` works differently for vectors and an S3 object with class = factor.* 

5.  Create a factor called `team` that has the values "Jim", "Sally", and "Ed" but has an extra level for a missing teammate "Sue". What is the order of the levels? Create a new factor `team2` where you permute the order of the levels in `team` such that they go "Sue", "Sally", "Ed", and "Jim". Run `table()` on both `team` and `team2` to make sure you did not change the count of the elements.

```{r}

team <- factor(c("Jim", "Sally", "Ed"), levels = c("Jim", "Sally", "Ed", "Sue"))
team2 <- factor(team, levels = c("Sue", "Sally", "Ed", "Jim"))

table(team)
table(team2)
```


___________________________

# Vectors (Ch 3) cont...

## More on factors

Picking up where we left off from last lecture:

```{r}

lmod <- lm(breaks ~ wool + tension, data = warpbreaks)
summary(lmod)

```
Formulae

```{r}
z <- ~ x + y
str(z)
attributes(z)

z <- u ~ x + y

as.formula("u ~ x + y")

z[[2]]
z[[3]]
z[[3]][[2]]
```

Reordering for different output

```{r}

wb2 <- warpbreaks

wb2$wool <- factor(wb2$wool, levels = c("B","A"))
wb2$tension <- factor(wb2$tension, levels = c("M","L","H"))

lmod2 <- lm(breaks ~ wool + tension, data = wb2)

summary(lmod2)
```

Looking at the model matrices:

```{r}
head(model.matrix(lmod))

head(model.matrix(lmod2))

```

Appending values into a factor:

```{r}
f1 <- factor(letters[1:3])
c(f1, "d")

levels(f1) <- letters[1:4]

c(f1, "d")

f1[[4]] <- "d"
f1

f1[[5]] <- "e"
f1

f1 <- factor(letters[1:3])
c(f1, factor("d"))

```

Ordered factors (ordinal categorical variables)

```{r}

possibleGrades <- c("F","D","C","B","A")

set.seed(5410)

classGrades <- sample(possibleGrades, 20, replace = T)

classGrades

```


```{r}
classGrades <- ordered(classGrades, levels = possibleGrades)

SAT.scores <- as.numeric(classGrades) * 25 + runif(length(classGrades), 500, 700)
```


```{r}
tapply(SAT.scores, classGrades, mean)
```


```{r}
summary(MASS::polr(classGrades ~ SAT.scores))
```
## Dates

Dates build on the double type

```{r}
today <- Sys.Date()
typeof(today)
attributes(today)
```

```{r}
date <- as.Date("1970-02-01")
unclass(date)
```

_________________________________________

# History

Everything I ran today in class:

```{r}
# x <- c(4.2, pi, 2/3, 3+2i)
# y <- as.integer(x)
# y
# dim(y) <- c(2,2)
# y
# dim(y) <- NULL
# y
# attr(y, "dim") <- c(2,2)
# y
# y <- as.vector(y)
# y
# structure(y, dim = c(2,2))
# y
# y[[20]] <- 10
# y
# sum(y)
# sum(y, na.rm = T)
# z <- factor(rep(c("a","truck","100"),4))
# z
# attributes(z)
# w <- unclass(z)
# w
# class(w) <- "factor"
# w
# team <- factor(c("Jim", "Sally", "Ed"), levels = c("Jim", "Sally", "Ed", "Sue"))
# team
# table(team)
# str(team)
# attributes(team)
# team2 <- factor(team, levels = c("Sue", "Sally", "Ed", "Jim"))
# team <- factor(c("Jim", "Sally", "Ed"), levels = c("Jim", "Sally", "Ed", "Sue"))
# team2 <- factor(team, levels = c("Sue", "Sally", "Ed", "Jim"))
# table(team)
# table(team2)
# lmod <- lm(breaks ~ wool + tension, data = warpbreaks)
# summary(lmod)
# z <- ~ x + y
# str(z)
# attributes(z)
# z <- u ~ x + y
# z
# as.formula("u ~ x + y")
# z[[2]]
# z[[3]]
# z[[3]][[2]]
# z <- ~ x + y
# str(z)
# attributes(z)
# z <- u ~ x + y
# as.formula("u ~ x + y")
# z[[2]]
# z[[3]]
# z[[3]][[2]]
# wb2 <- warpbreaks
# wb2$wool <- factor(wb2$wool, levels = c("B","A"))
# wb2$tension <- factor(wb2$tension, levels = c("M","L","H"))
# lmod2 <- lm(breaks ~ wool + tension, data = wb2)
# summary(lmod2)
# head(model.matrix(lmod))
# head(model.matrix(lmod2))
# f1 <- factor(letters[1:3])
# f1
# c(f1, "d")
# levels(f1) <- letters[1:4]
# f1
# c(f1, "d")
# f1[[4]] <- "d"
# f1
# f1[[5]] <- "e"
# f1
# possibleGrades <- c("F","D","C","B","A")
# possibleGrades <- c("F","D","C","B","A")
# set.seed(5410)
# classGrades <- sample(possibleGrades, 20, replace = T)
# classGrades
# classGrades <- ordered(classGrades, levels = possibleGrades)
# SAT.scores <- as.numeric(classGrades) * 25 + runif(length(classGrades), 500, 700)
# tapply(SAT.scores, classGrades, mean)
# summary(MASS::polr(classGrades ~ SAT.scores))
# today <- Sys.Date()
# typeof(today)
# attributes(today)
# as.numeric(today)
# date <- as.Date("1970-02-01")
# unclass(date)
```


