---
title: "Lecture 3"
author: "Ben Ridenhour"
date: "2022-09-01"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
    include:
      in_header: "preamble.tex"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary

New topics for today were:

- Using the `lobstr` package
- From `lobstr` using the `obj_addr()`, `tracemem()`, and `ref()` functions to detect when R is copying objects
    + Copy-on-modify behavior of R
    + Modify-in-place behavior for environments
- Understanding when R creates new objects for various types: vectors, lists, data.frames, environments
- Briefly discussing indexing with `[[` versus `[`
- Garbage collection/memory usage

-----------------------------------

## Quiz

1.  Create a **single** vector that consists of:

    -   the 26 lower case letters of the alphabet,

    -   the numbers 1, 2, and 3 repeated each 10 times, and

    -   100 evenly spaced numbers between 3 and 17.
    
    ```{r}
    
    x <- c(letters[1:26],rep(1:3,each=10),seq(from=3,to=17,length=100))
    x
    
    ```

2.  What is the product $$\begin{bmatrix}1 & 2 & 3 \\ 4 & 5 & 6 \\ 7 & 8 & 9\end{bmatrix}\begin{bmatrix}2 \\ 4 \\ 6\end{bmatrix} = \qquad ?$$

```{r}
matrix(1:9,nrow=3,byrow = T) %*% c(2,4,6)
```

3.  What function do you use to get the storage type (e.g., integer) of a variable?

```{r}
typeof(x)
```

4.  Why does the following code produce an error?

    `day_temp <- data.frame(day = c("Mon", "Wed", "Fri"),`

    `temp = c(32, 40, 35, 28))`
    
    Elements are different lengths

5.  Explain why `c(T,T,F) && c(T,F,F)` returns a value of true.

```{r}
c(T,T,F) && c(T,F,F)
c(T,T,F) & c(T,F,F)
```

&& only looks at the first elements of the vectors to perform the logical test.

--------------------------

## Chapter 2

```{r}
library(lobstr)
```

```{r}
myvec <- 4:15
myvec2 <- myvec

obj_addr(myvec)
obj_addr(myvec2)

#copy-on-modify of R
myvec2[[7]] <- 100
obj_addr(myvec2)

```

```{r}

tracemem(myvec)

myvec2<-myvec

obj_addr(myvec) == obj_addr(myvec2)

myvec2[[1]] <- "a"

typeof(myvec2)
```



### Create a function

```{r}

myFunction <- function(x){
  x
}

myFunction(2)
myFunction("a")


myFunction(myvec) #no tracemem return

myFunction <- function(x, squareMe = F){
  if(squareMe) return(x^2)
  x
}

myFunction(myvec)
myFunction(myvec,squareMe=T)


```


### List vs Data Frame memory

Lists

```{r}
l1 <- list(1,2,3)
obj_addr(l1)
lapply(l1,obj_addr)

l2 <- l1

obj_addr(l1) == obj_addr(l2)

?ref
ref(l1)
ref(l1,l2)

l2[[3]] <- "a"
ref(l1,l2)

```


Data.frames

```{r}

d1 <- data.frame(x = 1:3, y = letters[1:3])
d2 <- d1
d3 <- d1

ref(d1,d2,d3)

d2[,1] <- d2[,1]*2

ref(d1,d2)

d3[2,] <- c(100,"z")
ref(d1,d3)

```

### Environments

```{r}

e1 <- rlang::env(a = 1, b = 2, c = 3)
e2 <- e1

e2$c
ref(e1,e2)

e2$c <- "bob"
e2$c

ref(e1,e2)
e1$c
```

### Memory management/garbage collection

```{r}
gc()
mem_used()
```

-----------------------------

## History

Here is the history from today's lecture:

```{r}
# x <- c(letters[1:26],rep(1:3,each=10),seq(from=3,to=17,length=100))
# x
# matrix(1:9,nrow=3,byrow = T)
# matrix(1:9,nrow=3,byrow = T) %*% c(2,4,6)
# typeof(x)
# c(T,T,F) && c(T,F,F)
# c(T,T,F) && c(T,F,F)
# c(T,T,F) & c(T,F,F)
# library(lobstr)
# install.packages("lobstr")
# library(lobstr)
# ?lobstr
# ??lobstr
# ls(2)
# ls(3)
# myvec <- 4:15
# myvec <- 4:15
# myvec2 <- myvec
# myvec <- 4:15
# myvec2 <- myvec
# obj_addr(myvec)
# obj_addr(myvec2)
# myvec2[[7]] <- 100
# myvec2
# obj_addr(myvec2)
# myvec <- 4:15
# myvec2 <- myvec
# obj_addr(myvec)
# obj_addr(myvec2)
# myvec2[[7]] <- 100
# obj_addr(myvec2)
# myvec2<-myvec
# obj_addr(myvec)
# obj_addr(myvec) == obj_addr(myvec2)
# tracemem(myvec)
# myvec2<-myvec
# obj_addr(myvec) == obj_addr(myvec2)
# myvec2[[1]] <- "a"
# typeof(myvec2)
# myFunction <- function(x){
# x
# }
# myFunction(2)
# myFunction("a")
# myFunction(myvec)
# myFunction <- function(x, squareMe = F){
# if(squareMe) return(x^2)
# x
# }
# myFunction(myvec)
# myFunction(myvec, squareMe = T)
# z<- myFunction(myvec, squareMe = T)
# z
# tracemem(myvec)
# myFunction(myvec)
# myFunction(myvec,squareMe=T)
# l1 <- list(1,2,3)
# l1
# l1[[2]]
# obj_addr(l1)
# lapply(l1,obj_addr)
# l2 <- l1
# obj_addr(l1) == obj_addr(l2)
# ?ref
# ref(l1)
# ref(l1,l2)
# l2[[3]] <- "a"
# ref(l1,l2)
# d1 <- data.frame(x = 1:3, y = letters[1:3])
# d1
# View(d1)
# d1 <- data.frame(x = 1:3, y = letters[1:3])
# d2 <- d1
# d3 <- d3
# d1 <- data.frame(x = 1:3, y = letters[1:3])
# d2 <- d1
# d3 <- d1
# ref(d1,d2,d3)
# d1[,1]
# d1[1,]
# d1[1,1]
# d1[1,1:2]
# d1[1:2,1:2]
# d2[,1] <- d2[,1]*2
# d2
# ref(d1,d2)
# d3[2,] <- c(100,"z")
# d3
# ref(d1,d3)
# e1 <- rlang::env(a = 1, b = 2, c = 3)
# e2 <- e1
# e2$c
# ref(e1,e2)
# e2$c <- "bob"
# e2$c
# ref(e1,e2)
# e1$c
# gc()
# gc()
# mem_used()

```