---
title: "Lecture 11"
author: "Ben Ridenhour"
date: "2022-10-06"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

# Summary

1. The quiz covered creating a more complex function.
    - You needed to use several default options, generate random values, run a linear model, and use some basic plotting
    - The "bonus" part of the quiz required understanding how to work vectors and flow-control to do a repetitive task
2. We continued on our discussion of functions. Topics were:
    - Scoping rules (name masking)
    - Lazy evaluation
    - Using dot-dot-dot (...) to pass arguments to subfunctions
    - Return values (implicit/explicit/invisible)
    - Halting execution

-----------------------------------------

# Quiz

Write a function that does the following (give everything other than `n` a default value of 1):

1.  Creates a set of random normal variates `x` of length `n` with mean `u` and standard deviation `s`.
2.  Using that set of random values, create the variable `y` that has the value `a*x+b`.
3.  Add an error term to `y` that is normal with mean 0 and standard deviation `s_err`.
4.  Runs an `lm()` of `y` versus `x` and then plots the points and the predicted line. NB: The primitive graphics function `abline()` can be used to quickly plot a bivariate regression. On the plot:
    -   Allow the user to specify the color of both the line and the points separately.
    -   Allow the user to specify the labels for the axes and the title
5.  What are the formals of the function? What attribute can you use to see the code for the function?
6.  BONUS: adjust your function so that it can repeat (and overlay) multiple iterations of 1-4 with potentially different values of any of the parameters.

```{r}

quiz <- function(n, u = 1, s = 1, a = 1, b = 1, s_err = 1, point_col = 1,
                 line_col = 1, xlabel = 1, ylabel = 1, title = 1){
  x <- rnorm(n, mean = u, sd = s)
  y <- a*x + b
  y <- y + rnorm(n, sd = s_err)
  mod <- lm(y ~ x)
  plot(x,y, xlab = xlabel, ylab = ylabel, main = title, col = point_col)
  abline(mod, col = line_col)
}

formals(quiz)
attr(quiz, "srcref")

```

Check our function:

```{r}
#defaults
quiz(20)
```

```{r}
quiz(50, u = 50, s = 20, a = 5, b = 100, 10, point_col = "red", line_col = "blue",
     xlabel = "Random X", ylabel = "Y", title = "5*x+y linear model")
```

-----------------------------------------

# Lecture Ch 6. Functions (cont.)

## Scoping

Think of scoping as the search path used to find saved values. R uses "name masking" which means that R uses the most "immediate" environment to find a value. In terms of immediacy the list goes:

1. in function (local environment)
2. in the environment of the function (`environment(f)`)
3. in global environment

Function names are treated differently than non-function names

```{r}
testEnv <- new.env()

testEnv$x <- 2
x <- "bob"

f <- function() print(x)

f()

environment(f) <- testEnv

f()

```
Bad names

```{r}
f2 <- function(x) x+2
f3 <- function(){
  f2 <- 10
  f2(f2)
}
f3()
```

## Lazy evaluation

R only ever evaluated **necessary** parts of function calls. For example:

```{r}
g <- function(x){
  "Hi"
}

g(stop("Danger, Will Robinson!"))

g()

```

Why does 

```
x_ok <- function(x){
  ! is.null(x) && length(x) == 1 && x > 0
}
```

work, while

```
x_bad <- function(x){
  ! is.null(x) & length(x) == 1 & x > 0
}
```

does not?

## Dot-dot-dot (...)

`...` is used to pass other unlisted arguments to a function. Typically, this is used to pass unspecified arguments to a function within a function.

```{r}
f <- function(x, ...){
  sum(x^2, ...)
}

y <- c(1:10, NA, 11)

f(y)

f(y, na.rm = T)
```

`...` can be converted to a `list` and then used:

```{r}
f <- function(...) list(...)

f(x=1:4, y=c("ab","cd"), z = 3 + 4i)

g <- function(...){
  otherArgs <- list(...)
  exp(otherArgs$x)
}

g(x=1:4, y=c("ab","cd"), z = 3 + 4i)

```
Less commonly, we can refer to ... arguments by position with ..[index]

```{r}

g <- function(...){
  exp(..1)
  
}

g(x=1:4, y=c("ab","cd"), z = 3 + 4i)

```

## Return values

We can return values in 3 different ways:

1. implicit: `f <- function(x) x^3`
2. explicit: `f <- function(x) return(x^3)`
3. invisibly: `f <- function(x) invisible(x^3)`

```{r}
f <- function(x) invisible(x^3)

f(2)

val <- f(2)
val

withVisible(f(2))
```

## Halting evaluation

`stop()` can be used to halt evaluation and return and error.

```{r}
f <- function(x){
  stop("None shall pass")
  x^3
}

f(5)
```


-----------------------------------------

# History


```{r}
# Hmmmm... I lost today's history.
# Sadness.
```




