---
title: "Lecture 12"
author: "Ben Ridenhour"
date: "2022-10-11"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

# Summary

1. The quiz covered several aspects of functions including: ways of calling functions, scoping, return values, use of `...`, and stopping function execution.

2. We finished covering Ch. 6 for the lecture. We covered using `on.exit`, declaring prefix vs. infix functions, and declaring replacement functions

3. Toward the end of class, groups worked together on the "bonus" question on the quiz from last Thursday (6-Oct). I asked everyone to complete this task **before** handing in the assignment.

4. I forgot to turn on the Zoom recording until after I discussed the quiz. Sorry!

--------------------------------

# Quiz 

1.  Suppose you want calculate $e^{\sum_{i=1}^n x_i}$. What are 3 different ways to code this quantity?

```{r}
x <- 1:5

y <- sum(x); exp(y) #intermediate storage

exp(sum(x)) #nesting functions

library(magrittr)

x %>% sum() %>% exp()

#### BONUS 4th

x |> sum() |> exp()
```

2.  What is the order of scoping in R functions? (In other words, in what order does R check environments for the existence of a variable; we talked about 3 environments in class.)

```{r}
# in function

f <- function(){x <- 3; x}
f()


# global environment
f <- function() x
f()

# in the functions environment
environment(f)

theEnv <- new.env()
environment(f) <- theEnv
theEnv$x <- "bob"
f()
```


3.  What kind of return value does the assignment operator `<-` have? What function can help us check this type of return value?

```{r}

#we get an invisible return with <-
a <- 2

withVisible(a <- 2)
```


4.  How do we anonymously (without specifying them by name) pass arguments from a top level function to sub-functions? What are 3 ways to pass these arguments to subfunctions?

```{r}

#we use ... to pass off anonymous arguments

f <- function(...){
  sum(...) #no manipulation
}


f <- function(...){
  argList <- list(...)
  sum(argList[["x"]]) #this assumes that ... has an element named x
}

f <- function(...){
  sum(..1) #take the first argument in ... and pass it to sum
}


```

5.  Write a simple function that takes a single argument, performs a logical check of the argument, and finally stops executions with an **informative** (!) error message for the user if the logical test fails.

```{r}

f <- function(x){
  if(! is.integer(x)) stop("x was not an integer!")
  x^2
}

f(2L)

f(2.1)
```

--------------------------------

# Lecture Ch 6. Functions continued...

## Exiting gracefully

We can use the `on.exit( expr, add = T)` command to have functions do something when they stop. `on.exit` commands run in the order they appear in the function body unless we use the `after = F` option.

```{r}

f <- function(x){
  on.exit(message("a"), add = T)
  on.exit(message("b"), add = T)
  log(x)
}

f(10)

f <- function(x){
  on.exit(message("a"), add = T)
  on.exit(message("b"), add = T, after = F)
  log(x)
}

f(10)

```


## 3 type of functions

### Infix vs prefix functions

Most of the functions that you write are prefix. These have the structure of `FUNCTION(ARGS)`. We've also seen several "infix" functions, where the function goes in the middle of the arguments and evaluation proceeds from left to right.

For example,

```{r}
3 + 4

a <- 3

#we can always change an infix operator to be a prefix function

`+`(3,4)


```

We can write our own infix functions if we wish! To do this we use the `% SYMB %` construction.

```{r}

`%omg%` <- function(LHS, RHS) LHS^RHS

4 %omg% 3


```
### Replacement Functions

R has numerous functions that modify arguments "in place" and these are called replacement functions.

```{r}
x <- 1:3
names(x) <-  letters[1:3] #replacing the names attribute with something new
```

In order to write a replacement function, we must use arguments named `x` and `value` and write the function in this manner:

```{r}
`setBob<-` <- function(x, value){
  attr(x, "Bob") <- value
  x
}

z <- 1:5

setBob(z) <- "ooph"

attributes(z)

```
If you need more arguments for your replacement function, put those arguments **between** `x` and `value`.

```{r}
`setBob<-` <- function(x, count = 1,  value){
  attr(x, "Bob") <- rep(value, count)
  x
}

setBob(z) <- "yes"

attributes(z)

setBob(z,3) <- "no" #set count = 3

attributes(z)

```
--------------------------------

# History


Commands I ran today...

```{r}
# x <- 1:5
# y <- sum(x); exp(y) #intermediate storage
# exp(sum(x)) #nesting functions
# library(magrittr)
# x %>% sum() %>% exp()
# x <- 1:5
# y <- sum(x); exp(y) #intermediate storage
# exp(sum(x)) #nesting functions
# library(magrittr)
# x %>% sum() %>% exp()
# #### BONUS 4th
# x |> sum() |> exp()
# # in function
# f <- function(){x <- 3; x}
# f()
# # global environment
# f <- function() x
# f()
# # in the functions environment
# environment(f)
# theEnv <- new.env()
# environment(f) <- theEnv
# theEnv$x <- "bob"
# f()
# a <- 2
# #we get an invisible return with <-
# a <- 2
# withVisible(a <- 2)
# f <- function(x){
# if(! is.integer(x)) stop("x was not an integer!")
# x^2
# }
# f(2L)
# f <- function(x){
# if(! is.integer(x)) stop("x was not an integer!")
# x^2
# }
# f(2L)
# f(2.1)
# f <- function(x){
# on.exit(message("a"), add = T)
# on.exit(message("b"), add = T)
# log(x)
# }
# f(10)
# f <- function(x){
# on.exit(message("a"), add = T)
# on.exit(message("b"), add = T)
# log(x)
# }
# f(10)
# f <- function(x){
# on.exit(message("a"), add = T)
# on.exit(message("b"), add = T, after = F)
# log(x)
# }
# f(10)
# 3 + 4
# a <- 3
# #we can always change an infix operator to be a prefix function
# `+`(3,4)
# `%omg%` <- function(LHS, RHS) LHS^RHS
# 4 %omg% 3
# x <- 1:3
# names(x) <-  letters[1:3] #replacing the names attribute with something new
# `setBob <-` <- function(x, value){
# attr(x, "Bob") <- value
# x
# }
# z <- 1:5
# setBob(z) <- "ooph"
# `setBob<-` <- function(x, value){
# attr(x, "Bob") <- value
# x
# }
# z <- 1:5
# setBob(z) <- "ooph"
# attributes(z)

```

