---
title: "Lecture 7"
author: "Ben Ridenhour"
date: "2022-09-15"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
    include:
      in_header: "preamble.tex"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

- Nothing new on the quiz coverage, but there was a `lobstr` "refresher" in the third question
    + We did mention usage of the `paste()` command
- Finished the discussion of Ch. 3 
    + Complex structures of tibbles and data frames (e.g., lists of lists)
    + Manipulation: `cbind()`, `rbind()`, erasing columns, `add_column()`, `$` operator
- Began Ch. 4, Subsetting
    + Mainly covered ways of subsetting vectors: selection of individual or multiple elements with integers (positive or negative) or logical values, blank `[]`, and zero length return values
    + Using names to select elements of, e.g., named vectors or matrices (row or column names for subsetting)
- **Exam next Thursday, 22 Sept. over RStudio/R usage (general usage) and Ch. 2-4. **

-----------------------------

# Quiz

1.  What is the time difference between on `14-Sep-2022 2:32 PM` in St. Kitts and `17-Sep-2022 3:45 AM` in London in seconds? When you create your data-time objects, pass the **exact** date strings in the question. To find the correct time zones, remember to use `OlsonNames()`.

```{r}
fmt <- "%d-%b-%Y %I:%M %p"

dt1 <- as.POSIXct("14-Sep-2022 2:32 PM", tz = "America/St_Kitts",
                  format = fmt)

dt2 <- as.POSIXct("17-Sep-2022 3:45 AM", tz = "Europe/London",
                  format = fmt)
  
difftime(dt2, dt1, units = "secs")

format(dt1, "%d-%B-%y")
```

2.  Other than having different generic functions (e.g., `print()`), what are two differences between data frames and tibbles that were mentioned in the last lecture?

   - Row names get dropped in tibbles, and we can create subsequent columns in tibbles using earlier specified columns
   
```{r}
library(tibble)
tibble(x = 1:10, y = x^2)
```

3.  Create a data frame that has 100 rows with a the first column being random Poisson variables with mean 5 and the second column being random binomial variables with size = 10 and probability = 0.5. After creating the data frame, name the rows R1 ... R100 (use the `paste()` command to manufacture the 100 labels). After creating the data frame do the following:

```{r}
df <- data.frame(pois_var = rpois(100,5), bin_var = rbinom(100, 10, 0.5))

rownames(df) <- paste("R",1:100,sep="")
```

    -   Convert the data frame to a tibble (be sure keep the row names!)
    
    ```{r}
    tb <- as_tibble(df, rownames = "MyRowNames")
    tb
    ```

    -   Compare the memory usage of the tibble of and the data frame
    
```{r}
library(lobstr)

obj_size(df)
obj_size(tb)

```
    
    -   Compare the memory storage locations of the tibble and data frame
    
```{r}
obj_addr(df)
obj_addr(tb)

ref(df, tb)
```

    -   What do the two comparisons above tell you about tibbles relative to data frames?
    
    Tibbles take more memory!
    
# Lecture 7

## Ch. 3 Tibbles and Data frames (cont.)

Data frames and tibbles can have complex structures. They can hold lists of lists.

```{r}

data.frame(x = 1:3, y = I(list(1:2, 1:3, 1:4)))

tibble(x = 1:3, y = list(1:2, 1:3, 1:4))

lm1 <- lm(mpg ~ cyl + hp, data = mtcars)

lm2 <- lm(mpg ~ cyl * hp, data = mtcars)

lm3 <- lm1 <- lm(mpg ~ cyl * hp - 1, data = mtcars)

data.frame(x = 1:3, mods = I(list(lm1, lm2, lm3)))

```


Some manipulation commands

```{r}
#erase a column

head(df)
df$pois_var <- NULL

#appending rows or columns
tb2 <- tibble(v1=1:3, v2=letters[1:3])

#row binding
rbind(tb2, c(1,"bob")) #forced our int v1 to a char -- bad!
rbind(tb2, list(1, "bob")) #forced int to double
rbind(tb2, list(1L, "bob"))

#column binding
str(cbind(tb2, v3=4:6)) #doing this forces tibble to data frame
str(add_column(tb2, v3=4:6))

#cbind and rbind work for data frame
df2 <- as.data.frame(tb2)
rbind(df2, list(1L, "bob"))
cbind(df2, v3 = 4:6)

tb2$newCol <- 7:9
df2$newCol <- 7:9
```

# Ch 4 Subsetting

## Atomic vectors

There are 6 (ish) ways to subset vectors. We've talked about `[`, `[[`, and `$`.

```{r}
x <- runif(5)
x
```

Using positive integers
```{r}
x[2]
x[[2]]

x[c(1,4)]
x[rep(3,3)]
```

Negative integers

```{r}
x[c(-3,-2)]
x[-3:-2]
x[-c(3,2)]
```


Logical values

```{r}
x[c(T,F)]
```

"Blank"

```{r}
x[] #returns everyting in the index
```

Zero

```{r}
x[0] #keeps the structure of the object but drops all elements
```


Named vectors

```{r}
y <- setNames(x, LETTERS[1:5])

y[c("A","E")]

y[c("F")]
```

Matrices
```{r}
x <- matrix(1:6, nrow = 2)
colnames(x) <- paste("C",1:3,sep="")
rownames(x) <- paste("R",1:2,sep="")
attributes(x)
dimnames(x)
x["R1","C2"]

```



_________________________________

# History

R commands from today's lecture:

```{r}
# fmt <- "%d-%b-%Y %I:%M %p"
# dt1 <- as.POSICct("14-Sep-2022 2:32 PM", tz = "America/St_Kitts",
# format = fmt)
# fmt <- "%d-%b-%Y %I:%M %p"
# dt1 <- as.POSIXct("14-Sep-2022 2:32 PM", tz = "America/St_Kitts",
# format = fmt)
# dt1
# fmt <- "%d-%b-%Y %I:%M %p"
# dt1 <- as.POSIXct("14-Sep-2022 2:32 PM", tz = "America/St_Kitts",
# format = fmt)
# dt2 <- as.POSIXct("17-Sep-2022 3:45 AM", tz = "Europe/London",
# format = fmt)
# difftime(dt2, dt1, units = "secs")
# format(dt1, "%d-%B-%y")
# library(tibble)
# tibble(x = 1:10, y = x^2)
# df <- data.frame(pois_var = rpois(100,5), bin_var = rbinom(100, 10, 0.5))
# head(df)
# paste("R",1:100, "a",sep=c("","-"))
# rownames(df) <- paste("R",1:100,sep="")
# head(df)
# tb <- as_tibble(df, rownames = "MyRowNames")
# tb
# library(lobstr)
# library(lobstr)
# obj_size(df)
# obj_size(tb)
# obj_addr(df)
# obj_addr(tb)
# ref(df, tb)
# data.frame(x = 1:3, y = I(list(1:2, 1:3, 1:4)))
# data.frame(x = 1:3, y = I(list(1:2, 1:3, 1:4)))
# data.frame(x = 1:3, y = I(list(1:2, 1:3, 1:4)))
# tibble(x = 1:3, y = list(1:2, 1:3, 1:4))
# tibble
# tibble(x = 1:3, y = list(1:2, 1:3, 1:4))
# head(mtcars)
# data.frame(x = 1:3, y = I(list(1:2, 1:3, 1:4)))
# tibble(x = 1:3, y = list(1:2, 1:3, 1:4))
# lm1 <- lm(mpg ~ cyl + hp, data = mtcars)
# lm2 <- lm(mpg ~ cyl * hp, data = mtcars)
# lm3 <- lm1 <- lm(mpg ~ cyl * hp - 1, data = mtcars)
# data.frame(x = 1:3, mods = I(list(lm1, lm2, lm3)))
# data.frame(x = 1:3, mods = I(list(lm1, lm2, lm3)))
# str(lm1)
# head(df)
# View(df)
# df$pois_var <- NULL
# head(df)
# #appending rows or columns
# tb2 <- tibble(v1=1:3, v2=letters[1:3])
# rbind(tb2, c(1, "bob"))
# rbind(tb2, list(1, "bob"))
# rbind(tb2, list(1, "bob"))
# rbind(tb2, list(1L, "bob"))
# rbind(tb2, list(1L, "bob"))
# str(cbind(tb2, v3=4:6))
# str(add_column(tb2, v3=4:6))
# #cbind and rbind work for data frame
# df2 <- as.data.frame(tb2)
# df2
# rbind(df2, list(1L, "bob"))
# str(rbind(df2, list(1L, "bob")))
# cbind(df2, v3 = 4:6)
# tb2$newCol <- 7:9
# tb2
# df2$newCol <- 7:9
# df2
# x <- runif(5)
# x
# x[2]
# x[[2]]
# x[c(1,4)]
# x[rep(3,3)]
# x[c(-3,-2)]
# x[-3:-2]
# x[-c(3,2)]
# x[c(T,F)]
# x[]
# x[0]
# y <- setNames(x, LETTERS[1:5])
# y
# attributes(y)
# y <- setNames(x, LETTERS[1:5])
# y[c("A","E")]
# y[c("F")]
# x <- matrix(1:6, nrow = 2)
# colnames(x) <- paste("C",1:3,sep="")
# rownames(x) <- paste("R",1:2,sep="")
# attributes(x)
# dimnames(x)
# x["R1","C1"]
# x["R1","C2"]

```