---
title: "Lecture 10"
author: "Ben Ridenhour"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
date: "2022-10-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

-------------------------------------------------

# Summary

- We covered several topic during the quiz, but the focus was on using subsetting and flow control
   + Creation of data storage outside of for-loops
   + Random sampling, in particular drawing unique subsets of data
   + Using `lm` and getting coefficients
   + Some data manipulation
   + As "non-quiz" material, we covered some plotting basics (e.g., the `plot()` and `points()` commands) and making predictions for `lm`
- We began our discussion of Ch 6. which cover use and creation of functions. We talked about:
   + The 3 parts of a functions
   + Anonymous (headless) functions
   + Passing arguments as lists
   + Different ways of running a series of commands


------------------------------------------------

# Quiz

1.  Create 50 **unique** subsets by generating unique subsets of indices and storing them. See the help for the `unique()` function as a method of guaranteeing all 50 sets are unique.

```{r}
dim(quakes)
x <- matrix(ncol=50,nrow=500)
for(i in 1:50) x[,i] <- sample(nrow(quakes), nrow(quakes)*0.5)
ncol(unique(x, MARGIN = 2)) == ncol(x)

```


2.  Run and store the coefficients of the linear model for each of the 50 subsets.

```{r}
res <- vector("list", ncol(x))
for(i in 1:ncol(x)){
  res[[i]] <- coef(lm(mag ~ (depth + long)^2, data = quakes[x[,i],]))
}

res <- as.data.frame(t(as.data.frame(res)))
rownames(res) <- NULL

```

3.  Calculate the mean and standard errors of the resulting parameters estimates (there are 4 parameters in the model).

```{r}

ests <- colMeans(res)
ests
apply(res, 2, var)^0.5

```

4.  Compare the means and standard errors to the theoretical estimates of the means and standard errors given by `lm()`.


```{r}

summary(mod <- lm(mag ~ depth*long, data=quakes))

```

Bonus examination:

```{r}
plot(quakes$depth, quakes$mag, xlab = "Depth", ylab = "Magnitude",
     main = "Earthquake dependence on depth") #base layer

predMag <- predict(mod, data.frame(depth = 40:680, long = mean(quakes$long)))

points(40:680, predMag, type = "l", col = "red", lwd = 2)

robustMag <- ests[['(Intercept)']] + ests[['long']] * mean(quakes$long) + ests[['depth']] * 40:680 + ests[['depth:long']] *  mean(quakes$long) * 40:680

points(40:680, robustMag, type = "l", col = "blue", lwd = 2, lty = 2)


```


------------------------------------------------

# Lecture 10 - Ch 6. Functions

Functions have 3 parts:
1. formals (signature, arguments)
2. body (expression to be evaluated at run time)
3. environment (where the function is executed/determines the search path for values)

Remember that functions have class "function" and type of (`typeof()`) closure.

## Getting information about functions

```{r}
testFn <- function(a,b){
  #simple function
  a * log(b)
}

formals(testFn)
args(testFn) 

body(testFn)
attr(testFn, "srcref") #shows more formatting etc than body()

environment(testFn)

```
## Anonymous functions

Anonymous functions are created temporarily and have no name. Most often we use these in functions that call a secondary function. For examply, we use the commonly in the `apply` family of commands, e.g.:

```{r}

x <- matrix(rpois(100,20), nrow = 10)

apply(x, 1, function(row) sum(log(row))) #sum of the logged values of each row
```
## Passing arguments as a list

Constructing argument lists can be handy (at times):

```{r}
argList <- list(1:10, na.rm=T)
do.call(mean, argList) #runs mean(1:10, na.rm = T)
```

## Running functions "in line" (infix functions)

In the library `magrittr` there is the infix function "and then" `%>%`.

These are all ways we can chain together functions:

1. `y <- f(x); g(y)` [intermediate storage]
2. `g(f(x))` [nesting functions]
3. `x %>% f() %>% g() -> saveMe`

```{r}
library(magrittr)

x <- 1:10

y <- sum(x); log(y)

log(sum(x))

x %>% sum() %>% log()


```
## Scoping and more...

To be continued!

------------------------------------------------

# History

```{r}
# View(quakes)
# knitr::opts_chunk$set(echo = TRUE)
# dim(quakes)
# x <- matrix(ncol=50,nrow=500)
# for(i in 1:50) x[,i] <- sample(nrow(quakes), nrow(quakes)*0.5)
# ncol(unique(x, MARGIN = 2)) == ncol(x)
# res <- vector("list", ncol(x))
# for(i in 1:ncol(x)){
# res[[i]] <- coef(lm(mag ~ (depth + long)^2, data = quakes[x[,i]]))
# }
# res <- vector("list", ncol(x))
# for(i in 1:ncol(x)){
# res[[i]] <- coef(lm(mag ~ (depth + long)^2, data = quakes[x[,i],]))
# }
# res
# res <- vector("list", ncol(x))
# for(i in 1:ncol(x)){
# res[[i]] <- coef(lm(mag ~ (depth + long)^2, data = quakes[x[,i],]))
# }
# res <- as.data.frame(t(as.data.frame(res)))
# rownames(res) <- NULL
# View(res)
# colMeans(res)
# apply(res, 2, var)^0.5
# summary(mod <- lm(mag ~ depth*long, data=quakes))
# plot(quakes$depth, quakes$mag, xlab = "Depth", ylab = "Magnitude",
# main = "Earthquake dependence on depth")
# range(quakes$depth)
# data.frame(depth = 40:680, long = mean(quakes$long))
# plot(quakes$depth, quakes$mag, xlab = "Depth", ylab = "Magnitude",
# main = "Earthquake dependence on depth") #base layer
# predMag <- predict(mod, data.frame(depth = 40:680, long = mean(quakes$long)))
# points(40:680, predMag, type = "l", col = "red", lwd = 2)
# ests <- colMeans(res)
# ests
# apply(res, 2, var)^0.5
# ests
# plot(quakes$depth, quakes$mag, xlab = "Depth", ylab = "Magnitude",
# main = "Earthquake dependence on depth") #base layer
# predMag <- predict(mod, data.frame(depth = 40:680, long = mean(quakes$long)))
# points(40:680, predMag, type = "l", col = "red", lwd = 2)
# robustMag <- ests[['(Intercept)']] + ests[['long']] * mean(quakes$long) + ests[['depth']] * 40:680 + ests[['depth:long']] *  mean(quakes$long) * 40:680
# points(40:680, robustMag, type = "l", col = "blue", lwd = 2, lty = 2)
# testFn <- function(a,b){
# #simple function
# a * log(b)
# }
# formals(testFn)
# args(testFn)
# body(testFn)
# attr(testFn, "srcref") #shows more formatting etc than body()
# environment(testFn)
# x <- matrix(rpois(100,20), nrow = 10)
# apply(x, 1, function(row) sum(log(row))) #sum of the logged values of each row
# argList <- list(1:10, na.rm=T)
# do.call(mean, argList) #runs mean(1:10, na.rm = T)
# library(magrittr)
# x <- 1:10
# y <- sum(x); log(y)
# log(sum(x))
# x %>% sum() %>% log()
```

