---
title: "Lecture 13"
author: "Ben Ridenhour"
date: "2022-10-18"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

# Summary

1. The quiz covered writing and using replacement and infix functions, testing for the existence of list elements, and high vs. low-level plotting commands.
2. We began Ch 7 on environments and discussed a number of functions (in `base` and `rlang`) for working with environments. We also discussed similarities and dissimilarities of lists and environments, which seem very similar on the surface.
3. We discussed super assignment (`<<-`) as well.

-----------------------------

# Quiz

1.  Write a replacement function that takes a vector and repeats it some number of times. Execute the function on the vector `1:5` and repeat it 3 times.

```{r}
`repMe<-` <- function(x, value) {x <-rep(x,value); x}
myVec <- 1:5
repMe(myVec) <- 3
myVec

```

2.  Write an infix function that does the same thing as (1). Execute the function.

```{r}
`%repMe%` <- function(LHS, RHS) rep(LHS, RHS)

1:5 %repMe% 3

```

3.  Adjust (1) so that it takes an optional argument and appends it to the vector prior to doing the repetition. Execute the command with arguments specifying the vector `1:5` repeated 3 times with an appended vector `6:10`.

```{r}
`repMe<-` <- function(x, append = NULL, value) {x <-rep(c(x,append),value); x}
myVec <- 1:5
repMe(myVec,6:10) <- 3
myVec
```

4.  Suppose you've been given a list and you don't know what items (elements) are in the list. What command can you use to return a logical value that tells you if the element `res` is in the list?

```{r}

someList <- list(a = 1:5, b = letters[1:3])

is.null(someList[["res"]])

```

5.  During our exercise last Thursday the involved plotting points and lines, several plotting commands were used/discussed. Which plotting command was a "high level" command? Name 2 commands that we used that are "primitive/low level" commands. Why is it important to know the difference? (How do high level and low level commands interact?)

`plot()` is high level. Primitive commands were `abline()` and `points()`. High level commands create new plots and low level plots add layers to high level commands.


-----------------------------

# Lecture Ch 7 Environments

Environments are similar (but different) to lists...

```{r}
x <- new.env() #create a new environment
x$a <- 3
x[["bob"]] <- "hi"
names(x)
```
We can also use the `env()` command in the `rlang` package.

```{r}
library(rlang)
y <- env(a = 2, b = "apple", c = 1:5)
names(y)
y$b
```
We can nest environments within environments:

```{r}
x$y <- y

#self referential
x$x <- x

parent.env(y) <- x
parent.env(y)

#rlang version
env_parent(y)
env_parents(y)
```
If we want the full list of parents:

```{r}
env_parents(y, last = empty_env())
```

Why does this fail?

```{r}
y$f <- function() a
y$f()
environment(y$f)
```
How do we fix it?

```{r}
environment(y$f) <- y
y$f()
```
```{r}
y[["a"]] <- NULL
y$f()
```
However:

```{r}
z <- x
z$a <- 4
x$a
```

Another way to see the contents of an environment is given by `rlang`:

```{r}
print(x)
env_print(x)
```
## Super assignment

"Super" assignment (`<<-`) assigns a value in the *parent* environment of a function **if it exists**, otherwise it creates/sets the variable in the *global* environment.

```{r}
y[["setA"]] <- function(val) A <<- val
environment(y[["setA"]]) <- y

y$setA("ack")

x$A <- "here"

y[["setA"]]("cool")
```

Another difference from lists is that we cannot use numeric indexing:

```{r}
y[[1]]
```
But we can check for existence like a list:

```{r}
y[["xyz"]]
```

Checking for equivalence:

```{r}
z <- x
#z == x
identical(z, x)
```

## Various rlang commands

```{r}
env_has(y, "a") #does a exist in y?
env_get(y, "a") #return a from y
#env_get(y, "xyz") #return an error because xyz doesn't exist
env_get(y, "xyz", default = "NotHere") #returns default if xyz is missing
env_poke(y, "f2", function(x) x^2) #set a value for f2 in y
env_bind(y, d=12, e="list") #this allows you to bind in multiple elements
env_unbind(y, "a")
```

-----------------------------

# History

```{r}
# 
# `repMe<-` <- function(x, value) {x <-rep(x,value); x}
# myVec <- 1:5
# repMe(myVec) <- 3
# myVec
# `%repMe%` <- function(LHS, RHS) rep(LHS, RHS)
# 1:5 %repMe% 3
# `repMe<-` <- function(x, append, value) {x <-rep(c(x,append),value); x}
# myVec <- 1:5
# repMe(myVec,6:10) <- 3
# myVec
# someList <- list(a = 1:5, b = letters[1:3])
# is.null(someList[["res"]])
# library(rlang)
# ?? rlang::.
# x <- new.env() #create a new environment
# x$a <- 3
# names(x)
# x <- new.env() #create a new environment
# x$a <- 3
# x[["bob"]] <- "hi"
# names(x)
# y <- env(a = 2, b = "apple", c = 1:5)
# names(y)
# y <- env(a = 2, b = "apple", c = 1:5)
# names(y)
# y$b
# z <- x
# z$a <- 4
# x$a
# print(x)
# env_print(x)
# View(y)
# x$y <- y
# #self referential
# x$x <- x
# View(x)
# x$y <- y
# #self referential
# x$x <- x
# parent.env(y) <- x
# parent.env(y)
# print(x)
# x$y <- y
# #self referential
# x$x <- x
# parent.env(y) <- x
# parent.env(y)
# #rlang version
# env_parent(y)
# env_parents(y)
# env_parents(y, last = empty_env())
# y$f <- function() a
# y$f()
# View(y)
# View(x)
# y$f <- function() a
# y$f()
# environment(y$f)
# environment(y$f) <- y
# y$f()
# View(y)
# View(x)
# y$a <- NULL
# y$f()
# y$f()
# View(y)
# View(y)
# y[["a"]] <- NULL
# y$f()
# y[["setA"]] <- function(val) A <<- val
# environment(y[["setA"]]) <- y
# y$setA("ack")
# y[["setA"]] <- function(val) A <<- val
# environment(y[["setA"]]) <- y
# y$setA("ack")
# x$A <- "here"
# y[["setA"]]("cool")
# View(x)
# y[[1]]
# y[["xyz"]]
# z <- x
# z <- x
# z <- x
# z == x
# z <- x
# #z == x
# identity(z, x)
# z <- x
# #z == x
# identical(z, x)
# env_has(y, "a") #does a exist in y?
# env_get(y, "a") #return a from y
# #env_get(y, "xyz") #return an error because xyz doesn't exist
# env_get(y, "xyz", default = "NotHere") #returns default if xyz is missing
# env_poke(y, "f2", function(x) x^2) #set a value for f2 in y
# env_bind(y, d=12, e="list") #this allows you to bind in multiple elements
# env_print(y)
# env_has(y, "a") #does a exist in y?
# env_get(y, "a") #return a from y
# #env_get(y, "xyz") #return an error because xyz doesn't exist
# env_get(y, "xyz", default = "NotHere") #returns default if xyz is missing
# env_poke(y, "f2", function(x) x^2) #set a value for f2 in y
# env_bind(y, d=12, e="list") #this allows you to bind in multiple elements
# env_unbind(y, "a")
# env_print(y)
# y$c <- NULL
# env_print(y)


```
