---
title: "Lecture 6"
author: "Ben Ridenhour"
date: "2022-09-13"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

- The quiz covered:
    + Using formulae, using the `update()` command on a formula, and the `terms.formula()` command
    + Appending values to a factor
    + Creating a toy data set using factors and `rnorm()`; using `tapply()`
    + Creating a date and performing simple algebra (+ or -) with them
- Formatting dates for input and output (help `strptime` for formatting flags)
- POSIXct date-times (run command `OlsonNames()` to see a list of possible time zones)
- Started discussing data frames versus tibbles (`library(tibble)`)
    + Discussion mainly focused on differences between a tibble and a data frame
    

------------------------------

# Quiz

1.  Suppose you have a dependent variable `y` and explanatory factors `A`, `B`, and `C`. What is **a** correct formula to model main effects of all 3 factors, including all 2 way interactions (e.g., A x B)? What if you just wanted the interaction between B and C? What about a formula that has the 3-way interaction A x B x C? Suppose you want to know the effect of log-temperature (all you have is measure of temperature in a variable called `temp`) on `y` with just main effects of the 3 factors; then what would the formula be?

```{r}
fm1 <- y ~ (A + B + C)^2
# ~ A + B + C + A:B + A:C + B:C alternate
fm2 <- y ~ A + B*C
# ~ A + B + C + B:C
fm3 <- y ~ (A + B + C)^3
update(fm1, . ~ . + A:B:C)
update(fm1, . ~ . - A:B)
terms.formula(fm1)
fm4 <- y ~ A + B + C + log(temp)
```

2.  Create a factor with values and levels of "cat", "dog", and "bird". Now append the values "lizard" and "horse" to the factor. Verify that the appended version is a factor with the expected 5 values and 5 levels.

```{r}
myFactor <- factor(c("cat","dog","bird"))
c(myFactor, c("lizard","horse"))
c(myFactor, factor("lizard","horse"))
myFactor <- factor(myFactor, levels = c("cat","dog","bird","lizard","horse"))
myFactor[[4]] <- "lizard"
myFactor[[5]] <- "horse"
myFactor
```

3.  Create a toy data set (as a data frame) that matches the following properties:

    -   Has 100 rows of data
    -   Has a variable that is a factor have a **random number** of "A", "B", "C", "D", or "E" values
    -   What is the count of each of the levels?
    -   Has a variable that is that has the following means per group but with a Normal variance of 2 (i.e. use `rnorm`): $\mu_A = 3, \mu_B = 5, \mu_C = 1, \mu_D = 4, \mu_E = 2$
    -   Use `tapply()` to verify both the means and variances of the variables

```{r}
var1 <- factor(sample(LETTERS[1:5], 100, replace=T))
table(var1)
var1 <- factor(var1, levels = c("C","E","A","D","B"))
var2 <- rnorm(100, as.integer(var1), sqrt(2))

tapply(var2, var1, mean)
tapply(var2, var1, var)

df <- data.frame(var1,var2)
```

4.  The Black Death (bubonic plague) first showed up in Dorset, England around 8 Aug 1348. What date was it 1000 days later? How many days have elapsed since the plague showed up in Dorset and today? Use the `difftime()` command to calculate the number of weeks in the elapsed time.

```{r}
oldDay <- as.Date("1348-08-08")

oldDay + 1000

Sys.Date() - oldDay

difftime(oldDay,Sys.Date(),units="weeks")
```

_______________________________

# Lecture 6: Vectors again (Ch 3 continued...)

## Formatting Dates

Continuing with dates... formatting. Look at the help for `strptime`.

```{r}
format(as.Date("1992-01-02"), "%m%y%d")
format(as.Date("1992-01-02"), "%d BOB %B")
as.Date("019202", format = "%m%y%d")

# Convert "Dec 1, 1993" to a date
as.Date("Dec 1, 1993", "%b %d, %Y")
```

POSIXct (date-time class)

```{r}

today <- as.POSIXct("2022-09-06 10:25", tz = "America/Los_Angeles")
typeof(today)
attributes(today)
```

See `? OlsonNames()` for time zones.

```{r}
OlsonNames() #print out the possible time zones for POSIXct
```

## Data frames and tibbles

```{r}
df1 <- data.frame(x = 1:3, y = LETTERS[1:3])
str(df1)
```

The book tells you that y will a factor by default. This is incorrect as of R 4.0.

```{r}
df2 <- data.frame(x = 1:3, y = LETTERS[1:3], stringsAsFactors = T)
str(df2)
```

Data frames can have many attributes:

rownames, colnames, names, nrow, ncol, length

```{r}
attr(df2, "names") <- c("var1", "var2")
df2
unlist(df2)
```
Tibbles are an updated version of data frames

```{r}
library(tibble)

t1 <- as.tibble(df2)
str(t1)
attributes(t1)
t1
```

```{r}

#special to tibbles, we can create columns based on previous columns

t2 <- tibble(v1 = 1:10, v2 = sample(v1))
t2
```

Tibbles hate rownames

```{r}

rownames(df2) <- paste("R",1:3,sep="")
as_tibble(df2) #dropped row names
as_tibble(df2, rownames = "names")
```


_____________________________________

# History

The complete list of commands used today...

```{r}
# fm1 <- y ~ (A + B + C)^2
# # ~ A + B + C + A:B + A:C + B:C alternate
# fm2 <- y ~ A + B*C
# # ~ A + B + C + B:C
# fm3 <- y ~ (A + B + C)^3
# update(fm1, . ~ . + A:B:C)
# update(fm1, . ~ . - A:B)
# terms.formula(f1)
# fm1 <- y ~ (A + B + C)^2
# # ~ A + B + C + A:B + A:C + B:C alternate
# fm2 <- y ~ A + B*C
# # ~ A + B + C + B:C
# fm3 <- y ~ (A + B + C)^3
# update(fm1, . ~ . + A:B:C)
# update(fm1, . ~ . - A:B)
# terms.formula(fm1)
# fm1 <- y ~ (A + B + C)^2
# # ~ A + B + C + A:B + A:C + B:C alternate
# fm2 <- y ~ A + B*C
# # ~ A + B + C + B:C
# fm3 <- y ~ (A + B + C)^3
# update(fm1, . ~ . + A:B:C)
# update(fm1, . ~ . - A:B)
# terms.formula(fm1)
# fm4 <- y ~ A + B + C + log(temp)
# myFactor <- factor(c("cat","dog","bird"))
# c(myFactor, c("lizard","horse"))
# c(myFactor, factor("lizard","horse"))
# myFactor <- factor(c("cat","dog","bird"))
# c(myFactor, c("lizard","horse"))
# c(myFactor, factor("lizard","horse"))
# myFactor <- factor(myFactor, levels = c("cat","dog","bird","lizard","horse"))
# myFactor
# myFactor <- factor(c("cat","dog","bird"))
# c(myFactor, c("lizard","horse"))
# c(myFactor, factor("lizard","horse"))
# myFactor <- factor(myFactor, levels = c("cat","dog","bird","lizard","horse"))
# myFactor[[4]] <- "lizard"
# myFactor[[5]] <- "horse"
# myFactor
# var1 <- factor(sample(100, LETTERS[1:5], replace=T))
# var1 <- factor(sample(LETTERS[1:5], 100, replace=T))
# table(var1)
# typeof(var1)
# attributes(var1)
# var1 <- factor(var1, levels = c("C","E","A","D","B"))
# as.integer(var1)
# str(var1)
# attributes(var1)
# unclass(var1)
# var2 <- rnorm(100, as.integer(var1), sqrt(2))
# var2
# tapply(var2, var1, mean)
# tapply(var2, var1, var)
# df <- data.frame(var1,var2)
# View(df)
# oldDay <- as.Date("1348-08-08")
# oldDay
# attributes(oldDay)
# oldDay + 1000
# Sys.Date() - oldDay
# difftime(oldDay,Sys.Date(),units="weeks")
# ?strptime
# format(as.Date("1992-01-02"), "%m%y%d")
# format(as.Data("1992-01-02"), "%d BOB %B")
# format(as.Date("1992-01-02"), "%d BOB %B")
# as.Date("010902", format = "%m%y%d")
# as.Date("019202", format = "%m%y%d")
# # Convert "Dec 1, 1993" to a date
# as.Date("Dec 1, 1993", "%b %d, %Y")
# today <- as.POSIXct("2022-09-06 10:25", tz = "America/Los_Angeles")
# typeof(today)
# attributes(today)
# unclass(today)
# ? OlsonNames()
# df1 <- data.frame(x = 1:3, y = LETTERS[1:3])
# df1
# str(df1)
# df2 <- data.frame(x = 1:3, y = LETTERS[1:3], stringsAsFactors = T)
# str(df2)
# attr(df2, "names") <- c("var1", "var2")
# df2
# attributes(df2)
# unlist(df2)
# attr(df2, "names") <- c("var1", "var2")
# df2
# unlist(df2)
# unlist(df1)
# library(tibble)
# t1 <- as.tibble(df2)
# library(tibble)
# t1 <- as.tibble(df2)
# str(t1)
# attributes(t1)
# library(tibble)
# t1 <- as.tibble(df2)
# str(t1)
# attributes(t1)
# t1
# dplyr::starwars
# #special to tibbles, we can create columns based on previous columns
# t2 <- tibble(v1 = 1:10, v2 = sample(v1))
# t2
# rownames(df2) <- paste("R",1:3,sep="")
# df2
# as_tibble(df2)
# as_tibble(df2, rownames = "names")

```




