---
title: "Lecture 14"
author: "Ben Ridenhour"
date: "2022-10-20"
output: 
  pdf_document:
      toc: true
      toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

# Summary

1. ** Exam 2 next Thursday, 27 Oct! **
2. We basically went over the quiz which covered a bunch of the functions for working with environments. I also had one question regarding the use of the `on.exit()` function.
3. We discussed recursive functions and wrote an example.
4. We began a discussion of the evaluation environment, call stacks, and debugging.


--------------------------------

# Quiz

1.  What is the only environment in R without a parent environment?

The empty environment has no parent; `empty_env()`.

2.  What package can we use that has many "convenience" functions for working with environments? (Load it!)

```{r}
library(rlang)
```

3.  Write a function that takes a single argument `n` and returns *n* random normals. When the function completes provide the user a `summary()` of the randoms. Also, give the user an informational message telling them how many random numbers were generated. Store the output and verify that is a vector of doubles.

```{r}
f <- function(n){
  on.exit(print(summary(x)))
  on.exit(message(paste("The function generated ",n, " random normals.")), add = T, after = F)
  x <- rnorm(n) 
}

test <- f(10)
test

```

4.  Create an environment `x` that has bindings `a = 3`, `b = "apple"`, and `c = 1:5`.

    -   Add 3 new bindings (say d, e, and f) using 3 different methods (functions).
    -   Print out a listing of all of the bindings using 2 different methods.
    -   Check if the environment has a binding named `bob` using 2 different methods.
    -   Write a function `f` and bind it to the environment. `f` should print out the value of `b` within the environment `x`.
    -   Return the value of `c` using an `rlang` function.
    -   Create a copy of `x` called `y`. Check if they are indeed "copies" of each other.
    -   Remove the value of `a` from `x`. Print out the contents of `y`.
    -   Print out the **entire** search path for the function `f`.


```{r}
# Create an environment `x` that has bindings `a = 3`, `b = "apple"`, and `c = 1:5`.

x <- env(a = 3, b = "apple", c = 1:5)
x <- new.env()
x$a <- 3
x$b <- "apple"
x[["c"]] <- 1:5

# Add 3 new bindings (say d, e, and f) using 3 different methods (functions).

x$d <- letters[1:3]
env_bind(x, e = "hi")
env_poke(x, "f", function(x) x^2)


# Print out a listing of all of the bindings using 2 different methods.

env_print(x)
names(x)

# Check if the environment has a binding named `bob` using 2 different methods.

x[["bob"]]
env_has(x, "bob")

# Write a function `f` and bind it to the environment. `f` should print out the value of `b` within the environment `x`

x$g <- function() b
environment(x$g) #global environment
environment(x$g) <- x
x$g()

# Return the value of `c` using an `rlang` function.

env_get(x, "c", default = NA)

# Create a copy of `x` called `y`. Check if they are indeed "copies" of each other.

y <- x
identical(x,y)

# Remove the value of `a` from `x`. Print out the contents of `y`.

env_unbind(x, "a")
env_print(y)


# Print out the **entire** search path for the function `f`.

environment(x$g)
env_parents(environment(x$g), last = empty_env())

```

--------------------------------

# Lecture Ch 7 Environments (cont.)

When working with environments, it can be useful to write recursive functions. By a "recursive" function, we mean a function that calls itself again/repeatedly. Let's create a function that searches the path for a particular binding:

```{r}
findIt <- function(name, env = current_env()){
  if(identical(env, empty_env())){
    stop("Can't find ", name)
  } else if(env_has(env, name)){
    return(env)
  } else {
    findIt(name, env_parent(env))
  }
}

```


```{r}
findIt("mean", x)
#findIt("huh", y)
findIt("glm")
```
```{r}
search() #base
search_envs() #rlang
```

## Evaluation environment

All variables created within a function exist in temporary environment we call the evaluation environment. 

```{r}
f <- function(){
  a <- 1
  print(a)
}
```

"a" is only created when f is evaluated and then disappears.

If we get an error we can use traceback() to get the **call stack**. 

```{r}
f <- function(x){x <- 2; g(x)}
g <- function(x){x <- 3; h(x)}
h <- function(x) stop("!!!")
f(1)
traceback()
```

The lobstr::cst() functions somewhat similarly:

```{r}
f <- function(x){x <- 2; g(x)}
g <- function(x){x <- 3; h(x)}
h <- function(x) lobstr::cst()
f(1)
```


Each step of stack is called a frame (evaluation context). Frames have 3 parts:

1. the expression
2. the evaluation environment
3. the parent environment


--------------------------------

# History

```{r}
# library(rlang)
# f <- function(n){
# on.exit(print(summary(x)))
# on.exit(message(paste("The function generated ",n, " random normals.")), add = T)
# x <- rnorm(n)
# }
# test <- f(10)
# test
# f <- function(n){
# on.exit(print(summary(x)))
# on.exit(message(paste("The function generated ",n, " random normals.")), add = T, after = F)
# x <- rnorm(n)
# }
# test <- f(10)
# test
# # Create an environment `x` that has bindings `a = 3`, `b = "apple"`, and `c = 1:5`.
# x <- env(a = 3, b = "apple", c = 1:5)
# x <- new.env()
# x$a <- 3
# x$b <- "apple"
# x[["c"]] <- 1:5
# # Add 3 new bindings (say d, e, and f) using 3 different methods (functions).
# x$d <- letters[1:3]
# env_bind(x, e = "hi")
# env_poke(x, f = function(x) x^2)
# # Create an environment `x` that has bindings `a = 3`, `b = "apple"`, and `c = 1:5`.
# x <- env(a = 3, b = "apple", c = 1:5)
# x <- new.env()
# x$a <- 3
# x$b <- "apple"
# x[["c"]] <- 1:5
# # Add 3 new bindings (say d, e, and f) using 3 different methods (functions).
# x$d <- letters[1:3]
# env_bind(x, e = "hi")
# env_poke(x, "f", function(x) x^2)
# # Print out a listing of all of the bindings using 2 different methods.
# env_print(x)
# names(x)
# # Check if the environment has a binding named `bob` using 2 different methods.
# x[["bob"]]
# env_has(x, "bob")
# # Write a function `f` and bind it to the environment. `f` should print out the value of `b` within the environment `x`
# x$g <- function() b
# environment(x$g) #global environment
# environment(x$g) <- x
# x$g()
# # Return the value of `c` using an `rlang` function.
# env_get(x, "c", default = NA)
# # Create a copy of `x` called `y`. Check if they are indeed "copies" of each other.
# y <- x
# identical(x,y)
# # Remove the value of `a` from `x`. Print out the contents of `y`.
# env_unbind(x, "a")
# env_print(y)
# # Print out the **entire** search path for the function `f`.
# env_parents(environment(x$g), last = empty_env())
# # Create an environment `x` that has bindings `a = 3`, `b = "apple"`, and `c = 1:5`.
# x <- env(a = 3, b = "apple", c = 1:5)
# x <- new.env()
# x$a <- 3
# x$b <- "apple"
# x[["c"]] <- 1:5
# # Add 3 new bindings (say d, e, and f) using 3 different methods (functions).
# x$d <- letters[1:3]
# env_bind(x, e = "hi")
# env_poke(x, "f", function(x) x^2)
# # Print out a listing of all of the bindings using 2 different methods.
# env_print(x)
# names(x)
# # Check if the environment has a binding named `bob` using 2 different methods.
# x[["bob"]]
# env_has(x, "bob")
# # Write a function `f` and bind it to the environment. `f` should print out the value of `b` within the environment `x`
# x$g <- function() b
# environment(x$g) #global environment
# environment(x$g) <- x
# x$g()
# # Return the value of `c` using an `rlang` function.
# env_get(x, "c", default = NA)
# # Create a copy of `x` called `y`. Check if they are indeed "copies" of each other.
# y <- x
# identical(x,y)
# # Remove the value of `a` from `x`. Print out the contents of `y`.
# env_unbind(x, "a")
# env_print(y)
# # Print out the **entire** search path for the function `f`.
# environment(x$g)
# env_parents(environment(x$g), last = empty_env())
# findIt <- function(name, env){
# if(identical(env, empty_env())){
# stop("Can't find", name)
# } else if(env_has(env, name)){
# return(env)
# } else {
# findIt(name, env_parent(env))
# }
# }
# findIt("mean", environment(x))
# environment(x)
# x
# findIt("mean",x)
# findIt("mean", x)
# fintIt("huh", y)
# findIt("mean", x)
# findIt("huh", y)
# findIt <- function(name, env){
# if(identical(env, empty_env())){
# stop("Can't find ", name)
# } else if(env_has(env, name)){
# return(env)
# } else {
# findIt(name, env_parent(env))
# }
# }
# findIt("mean", x)
# findIt("huh", y)
# findIt("mean", x)
# findIt("huh", y)
# findIt("mean", x)
# #findIt("huh", y)
# findIt("glm", current_env())
# findIt <- function(name, env = current_env()){
# if(identical(env, empty_env())){
# stop("Can't find ", name)
# } else if(env_has(env, name)){
# return(env)
# } else {
# findIt(name, env_parent(env))
# }
# }
# findIt("mean", x)
# #findIt("huh", y)
# findIt("glm")
# search()
# search_envs()
# f <- function(){
# a <- 1
# print(a)
# }
# f()
# f <- function(x){x <- 2; g(x)}
# g <- function(x){x <- 3; h(x)}
# h <- function(x) stop("!!!")
# f(1)
# f <- function(x){x <- 2; g(x)}
# g <- function(x){x <- 3; h(x)}
# h <- function(x) stop("!!!")
# f(1)
# traceback()

```