---
title: "Lecture 8"
author: "Ben Ridenhour"
date: "2022-09-20"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

- The quiz covered:
    + Manipulation of tibbles (and data frames) with addition/deletion columns and rows
    + Subsetting/slicing vectors using integer, negative integers, names, and logicals
    + Return 0 length vectors or full length vectors
    + Subsetting/slicing matrices to get columns, rows, or individual elements
- We continued Ch.4 on subsetting and covered:
    + Getting multiple (scattered) elements from a matrix (or data frame/tibble)
    + Using the `drop = F` flag for data frames to maintain structure
    + Use of `$` vs `[[` vs `[` (and briefly `@` or `slot()`)
    + Partial matching of `$`
    + Using named vectors to create a lookup
    + Permutation, unique sub-samples, bootstrapping
    + Selection using logical tests
- We talked about the exam **on Thursday**
    + Get the exam preview on Canvas for details


_________________

# Quiz

1.  Create a tibble that has 3 columns and 10 rows: the first column should be a factor, the second column should be a random normal, and the third column is the absolute value of the second column. Name the columns. After creating the tibble:
    -   Add a fourth column that is the natural logarithm of the third column. **Make sure your tibble remains a tibble.**
    -   Add a row with new (appropriate) values for the 4 variables.
    -   Delete the second column.
    -   Depending on how you added the fourth column, use an alternative method to bind in another column (data of your choosing).
    -   Coerce your tibble to a data frame and the check the object size of the two objects. Which is larger?
    
```{r}
library(tibble)
library(lobstr)

ti <- tibble(rf = factor(sample(c("a", "b", "c"),10, replace = T)),
             rn = rnorm(10),
             abrn = abs(rn))

ti <- add_column(ti, lgrn = log(ti$abrn))

#ti$lgrn <- log(ti$abrn)
#ti[[4]] <- log(ti$abrn)

ti <- rbind(ti, list("c", 0.2, 0.2, log(0.2)))
ti$rn <- NULL
#ti[[2]] <- NULL
#ti[,2] <- NULL

ti[[4]] <- rpois(11, 5)

df <- as.data.frame(ti)

obj_size(df)
obj_size(ti)
```

## Quiz (cont.) {.smaller}

2.  Create a vector with 100 random uniform values in the range of 0 to 5; before generating the random numbers, initialize the random number generator using the value 12. Do the following:
    -   Name the elements of the vector by pasting the letter "e" together with the results of `sample(1:100)`
    -   Drop every third element and find the sum of the resulting vector
    -   Keep elements 1-20 and 80-90 then find the mean of the resulting vector
    -   Return the elements e1-e10 as a vector
    -   Return an empty vector
    -   Return all the values of the vector while using the `[` operator
    -   Convert the vector to a 10 x 10 matrix. Get columns 5-7. Get rows 7-9. Get the element in the 7th row and 9th column.
    
```{r}

set.seed(12)
rv <- runif(100, max = 5)
names(rv) <- paste("e",sample(1:100),sep="")
sum(rv[c(T,T,F)])
sum(rv[-seq(3,100,by=3)])
#could also do with positive integers 1,2,4,5,7,8...

mean(rv[c(1:20, 80:90)])
rv[paste("e",1:10,sep="")]

rv[0]
rv[]

mat <- matrix(rv, nrow = 10)

mat[,5:7]
mat[7:9,]
mat[7,9]

```
    
_________________

# Lecture - Subsetting (Ch 4 cont.)


## Getting multiple elements that are scattered

```{r}

getThese <- matrix(ncol = 2, byrow = T,
                   c(1,1,
                     4,7,
                     7,9)
                   )

mat[getThese]

```

## Data frames

Just like matrices, the first index indicates rows and the second index indicates columns. We can pick or drop using integers, names, logicals, etc...

```{r}

df[,1]
typeof(df[,1])
class(df[,1])
str(df[,1])

df[,1,drop=F]
str(df[,1,drop=F])
```

In general, `[[` this gets a single item while `[` shortens the object

`$` is similar to `[[`

```{r}
df$rf
df[["rf"]]

z<- "rf"
df$z
df[[z]]
```

`$` does partial matching

```{r}
l1 <- list("bob" = 1:4, "bonnie" = 5:8)

l1$b #fails because bob and bonnie both start with "b"
l1$bo # fails because both start with "bo"
l1$bon #partial match succeeds

```

In S4, there is the `@` operator and `slot()` function that work like `$` and `[[`, respectively.

Subsetting using "" (nothing) is useful because it preserves structure

```{r}

mt1 <- mtcars

mt1[] <- lapply(mt1, as.integer)

is.data.frame(mt1)

```


```{r}

mt2 <- mtcars

mt2 <- lapply(mt2, as.integer)

is.data.frame(mt2)

mt2 <- as.data.frame(mt2)

```

## Applications of subsetting

### Names vectors as lookup

```{r}
x <- sample(letters[1:3], 10, replace = T)
vals <- c("a" = "Mon", "b" = "Tues", "c" = "Weds")

vals[x]
```

### Permutation and bootstrapping

```{r}

#Permutation (shuffling elements)

df[sample(nrow(df)),]

#Unique subsample (cross-validation techniques)

df[sample(nrow(df),4),]

#Bootstrapping

df[sample(nrow(df), replace = T),]

```
### Logical subsetting

There are many types of logical tests that can all be used for subsetting

```{r}
mtcars[mtcars$gear == 5, ]
mtcars[mtcars$gear == 5 | mtcars$gear == 4,]
mtcars[mtcars$gear == 5 & mtcars$mpg > 30, ]
mtcars[mtcars$gear %in% 3:5, ]
```

-----------------

# History 

```{r}
# knitr::opts_chunk$set(echo = TRUE)
# factor(sample(c("a", "b", "c")),10))
# factor(sample(c("a", "b", "c"),10))
# factor(sample(c("a", "b", "c"),10, replace = T))
# ti <- tibble(rf = factor(sample(c("a", "b", "c"),10, replace = T)),
# rn = rnorm(10),
# abrn = abs(rn))
# library(tibble)
# library(lobstr)
# ti <- tibble(rf = factor(sample(c("a", "b", "c"),10, replace = T)),
# rn = rnorm(10),
# abrn = abs(rn))
# View(ti)
# t1 <- add_column(lgrn = log(t1$abrn))
# t1 <- add_column(t1, lgrn = log(t1$abrn) )
# ti <- add_column(ti, lgrn = log(ti$abrn) )
# ti
# ti <- rbind(ti, list("c", 0.2, 0.2, log(0.2)))
# str(ti)
# ti$rn <- NULL
# ti
# ti[[4]] <- rpois(11)
# ti[[4]] <- rpois(11, 5)
# ti
# df <- as.data.frame(ti)
# obj_size(df)
# obj_size(ti)
# rv <- runif(100, max = 5)
# rv
# set.seed(12)
# rv <- runif(100, max = 5)
# names(rv) <- paste("e",sample(1:100),sep="")
# sum(rv[c(T,T,F)])
# sum(rv[-seq(3,100,by=3)])
# #could also do with positive integers 1,2,4,5,7,8...
# mean(rv[c(1:20, 80:90)])
# rv[paste("e",1:10,sep="")]
# rv[0]
# rv[]
# mat <- matrix(rv, nrow = 10)
# mat[,5:7]
# mat[7:9,]
# mat[7,9]
# View(mat)
# getThese <- matrix(ncol = 2, byrow = T,
# c(1,1,
# 4,7,
# 7,9)
# )
# mat[getThese]
# df[,1]
# typeof(df[,1])
# class(df[,1])
# str(df[,1])
# df[,1,drop=F]
# df[,1,drop=F]
# str(df[,1,drop=F])
# df$rf
# df[["rf"]]
# z<- "rf"
# df$z
# df[[z]]
# l1 <- list("bob" = 1:4, "bonnie" = 5:8)
# l1$b
# l1$bo
# l1$bon
# l1$bo
# mt1 <- mtcars
# mt1[] <- lapply(mt1, as.integer)
# is.data.frame(mt1)
# mt1
# mt2 <- lapply(mt2, as.integer)
# mt2 <- mtcars
# mt2 <- lapply(mt2, as.integer)
# is.data.frame(mt2)
# str(mt2)
# mt2 <- as.data.frame(mt2)
# str(mt2)
# x <- sample(letters[1:3], 10, replace = T)
# vals <- c("a" = "Mon", "b" = "Tues", "c" = "Weds")
# vals[x]
# sample(11)
# nrow(df)
# df[sample(nrow(df)),]
# sample(nrow(df),4)
# sample(nrow(df),4)
# sample(nrow(df),4)
# sample(nrow(df),4)
# sample(nrow(df),4)
# df[sample(nrow(df),4),]
# sample(nrow(df), replace = T)
# df[sample(nrow(df), replace = T),]
# mtcars$gear == 5
# mtcars[mtcars$gear == 5, ]
# mtcars[mtcars$gear == 5 | mtcar$gear == 4,]
# mtcars[mtcars$gear == 5 | mtcars$gear == 4,]
# mtcars[mtcars$gear == 5 & mtcars$mpg > 30, ]
# mtcars[mtcars$gear %in% 3:5, ]

```

