---
title: "Lecture 2"
author: "Ben Ridenhour"
date: "2022-08-30"
output: 
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

+ We looked quickly at setting up an R project with Git support and RStudio's ability to work with other programming languages. 
+ We learned about the `typeof()` function
+ Covered some basic types: logical, integer, double, complex, character, function, and list
  - Other types are possible and will be talked about later
+ We discussed 3 different ways to create a vector: `c()`, `seq()`, and `rep()`
  - Showed some oddities of the way R does multiplication, addition, etc with vectors
  - Warning message about mismatched vector lengths 
+ Looked at creating matrices and arrays
  - Elementwise multiplication vs. matrix multiplication
+ Creating lists and using `lappy()`
  - `data.frames` as a special type of list where all elements of the list have the same length
+ Working with logical vectors and logic test
  - `==`, `!=`, `<`, `>`, `&`, `&&`, `|`, `||`


-------------------------------------------------

# Code from Class

## Lecture 2

```{r cars}
TRUE
FALSE

T 
F #abbreviated true false

typeof(1L) #L gives an integer type
typeof(1) #gives and double

typeof(3 + 2i)

typeof("hello")
typeof('world')

```

### Creating vectors

```{r}

x <- c(32, 2 + 4i, -51.5) #concatenate
y <- c(32, "a", 90)

z <- 5:10 #sequence from 5 to 10 

LETTERS[1:10] #upper case letters
letters[11:20] #lower case letters 

s1 <- seq(1,10)
s2 <- seq(1,10, by = 0.33)
s3 <- seq(1,10, le = 100)

r1 <- rep(1:4,5)
r2 <- rep(1:4,each=5)
r3 <- rep(1:4,length = 10)

```

### Tricky math?

```{r}

rep(1,20) * 1:3
rep(1,20) - 1:3


```


## More complex data types

```{r}
matrix(1,ncol=4,nrow=4)

matrix(1,4,4) * 1:4

matrix(1:4,4,4, byrow = T)

# matrix(1:4,4,4) %*% 1:3 #fails due to non-conformable arguments in matrix multiplication

matrix(1:4,4,4) %*% c(1,1,1,1)

```

 Here is an array:
```{r}
a1 <- array(rep(1:3,each=4), dim = c(2,2,3)) #dimensions come in as vector
a1
```

### Lists

```{r}

bob <- list(
  w = T,
  x = LETTERS[1:5],
  y = c(4, 9, 21),
  z = 10:30
)

bob$x #dollar sign operate to get element named x
bob[[3]] #get element by position in list

tom <- list(3,"a", 2+4i)

typeof(bob)

lapply(bob,typeof)
```
## Data frames

A special list where every element has the same length

```{r}
df1 <- data.frame(
  var1 = c("a","b","c"),
  var2 = rep(4,3),
  var3 = c(5,1,2)
)
  
```


### Logical variables

```{r}
#we can perform unary operations on logical variables

T + F
T + T
T * F
F - F

sum(runif(100) < 0.5)

5 == 5 #equivalence
5 != 5 #not equal

T & T #elementwise "and" comparison
T & F
F & F 

T | T #elementwise "or" comparison
T | F 
F | F 

l1 <- c(T, T, F)
l2 <- c(T, F, F)

l1 & l2
l1 | l2

l1 && l2 #first element "and" check
l1 || l2 #first element "or"

all(l1)
any(l1)
```


------------------------------

# History

The full history from today's lecture. You can use the `Code -> Comment/Uncomment Lines` to quickly change lines to being uncommented.

```{r}
# ? typeof
# help("typeof")
# typeof(54L)
# knitr::opts_chunk$set(echo = TRUE)
# typeof(3 + 2i)
# typeof("hello")
# typeof('world')
# typeof(`+`)
# typeof(+)
# typeof(ls)
# ls()
# x <- c(32, 2 + 4i, -51.5) #concatenate
# typeof(x)
# y <- c(32, "a", 90)
# typeof(y)
# z <- 5:10
# z
# LETTERS[1:10]
# letters[11:20]
# seq(1, 10)
# 1:10
# s1 <- seq(1,10)
# s1 <- seq(1,10, by = 0.33)
# s1 <- seq(1,10)
# s2 <- seq(1,10, by = 0.33)
# s2
# s3 <- seq(1,10, le = 100)
# s3
# lenght(s3)
# length(s3)
# r1 <- rep(1:4,5)
# r1
# r2 <- rep(1:4,each=5)
# r2
# r3 <- rep(1:4,length = 10)
# r3
# rep(1,20) * 1:3
# rep(1,20) - 1:3
# rep(1,20) * c(3, 5, 10)
# matrix(1,ncol=4,nrow=4)
# matrix(1,4,4) * 1:3
# matrix(1,4,4) * 1:4
# matrix(1:4,4,4)
# matrix(1:4,4,4, byrow = T)
# matrix(1:4,4,4)
# matrix(1:4,4,4) %*% 1:3
# matrix(1:4,4,4) %*% c(1,1,1,1)
# rep(1:3,each=4)
# a1 <- array(rep(1:3,each=4), dim = c(2,2,3)) #dimensions come in as vector
# a1
# array(rep(1:3,each=4), dim = c(2,3,3))
# a1 <- array(rep(1:3,each=4), dim = c(2,2,3)) #dimensions come in as vector
# a1
# bob <- list(
# w = T,
# x = LETTERS[1:5],
# y = c(4, 9, 21),
# z = 10:30
# )
# bob
# bob$x #dollar sign operate to get element named x
# bob[[3]]
# bob$y
# tom <- list(3,"a", 2+4i)
# tom
# typeof(bob)
# typeof(bob)
# lapply(bob,typeof)
# df1 <- data.frame(
# var1 = c("a","b","c"),
# var2 = rep(4,4),
# var3 = c(5,1,2)
# )
# df1 <- data.frame(
# var1 = c("a","b","c"),
# var2 = rep(4,3),
# var3 = c(5,1,2)
# )
# View(df1)
# T + F
# T + T
# T * F
# F - F
# runif(100)
# runif(100) < 0.5
# sum(runif(100) < 0.5)
# 5 == 5
# 5 != 5
# T & T
# T & F
# F & F
# T | T #elementwise "or" comparison
# T | F
# F | F
# l1 <- c(T, T, F)
# l2 <- c(T, F, F)
# l1 & l2
# l1 | l2
# l1 && l2
# l1 || l2 #first element "or"
# all(l1)
# l1
# any(l1)
```

