---
title: "Lecture 1 Summary"
author: "Ben Ridenhour"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Summary

Here is a brief list of the topics we covered today:

+ Syllabus (grading policy, book, etc)
  - Note that the book is available at <https://adv-r.hadley.nz> if do not wish to purchase a copy
+ Using RStudio
  - Usage of the 4 panes (Source, Environment, Console, Files/Help)
  - Going back with history
  - Running lines of scripts
  - And more...
+ R Markdown
  - Cheat sheet under the "Help > Cheat sheets" menu
  - YAML header and switching between knitting PDF, HTML, Word, and Notebook
  - Basic text input and using LaTeX
  - Entering R code in blocks
+ GitHub/R Project
  - The address for the course GitHub is <https://git.com/magbenji/Stat427Fall2022>
  - We will go over making an R project and GitHub again next week (because of my laptop's fail)
+ Some basic R
  - There is a list of basic R vocabulary at <http://adv-r.had.co.nz/Vocabulary.html>
  - How to read get help and read R help pages
  - Function signatures
  - Functions vs. operator (e.g., `assign()` vs. `<-`)
  - Installing, loading, and unloading packages (`install.packages()`, `library()`, and `unloadNamespace()`)
  - Scope, environment layers, and the `::` or `:::` operators
  
**Note: We will be working on Chapter 2 of the book next class period.**

And here is some gratuitous R code:

```{r BlockName}

set.seed(1234) #set the random seed
x <- runif(15) #create 15 random values from a [0,1] continuous uniform distribution
x #print out the values to see them
hist(x,10) #create a histogram of the values with 10 bins

```
  