---
title: "Lecture 15"
author: "Ben Ridenhour"
date: "2022-10-25"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

# Summary

1. The quiz covered using super assignment within nested environments, recursive functions, examining the call stack, and the concept of the evaluation environment.

2. The first part of lecture covered some topics from Ch 7 that we hadn't yet discussed: lazy bindings, active bindings, and namespaces. These were more FYI topics than ones I expect you to be able to use.

3. We began our discussion of Ch 8 on "conditions." We covered the basic idea of the condition system and the "signalling" side of the system. In particular, we focused on the the functions in `base` and `rlang` packages: `stop()`, `abort()`, `warning()`, `warn()`, and `message()`.

4. We also went over the exam preview (now available on Canvas).


--------------------------------

# Quiz

1. Create an environment `x`. Within `x` create a second environment `y`. In `y` bind a variable named `z` and give it a value of "hi". Create a function that exists in `y`. This function should do 2 things: 1) read the value of `z` in `y` and 2) uses super assignment to bind a variable named `w` in `x` that takes the current value of `z`.

```{r}
library(rlang)
x <- env(y = env(z = "hi", f = function() w <<- z))
parent.env(x$y) <- x
environment(x$y$f) <- x$y
#we need to have w in x for super assignment to write w in x (otherwise it
#it will be in the global environment)
env_bind(x, w = NULL)
x$y$f()
```

2. What is a recursive function? Write a short example of a recursive function. 

A recursive function is one that calls itself (repeatedly).
```{r}
#f <- function() f()
```

3. What 2 commands did we discuss to show the "call stack"? Create a set of 3 nested functions, the last of which throws an error. Use one of the commands to show the call stack after the error is thrown.

`traceback()` and `lobstr::cst()` can be used to see the call stack.

```{r}
f <- function(x) g(x)
g <- function(y) h(y)
h <- function(z) log(z)

f("a")

traceback()

f <- function(x) g(x)
g <- function(y) h(y)
h <- function(z) {lobstr::cst();log(z)}

f("a")
```

4. What is the "evaluation environment"?

The evaluation environment is the temporary environment created when we call a function.

----------------------------------


# Lecture Ch 8 Conditions (some last bits of chapter 7)

## Ch 7 oddities

In environments, you have "lazy" or "active" bindings. Lazy bindings in environments do not get evaluated until accessed:

```{r}
env_bind_lazy(x, wait = {Sys.sleep(1); 1})

system.time(print(x$wait))

system.time(print(x$wait))
```
Active bindings get evaluated every time they are accessed:

```{r}
env_bind_active(x, now = function() runif(1))
x$now
x$now
```
Namespaces are important environments that have some special features. The most notable is that they have a specified search path for bindings.

## Ch 8 Condtions

The "condition" system provides ways to signal something unusual or problematic might be happening and a way to deal with these issues. We can broadly divide the condition system into two parts: signalling and handling.

Signalling functions:
1. `stop()`
2. `warning()`
3. `message()`

Handling functions:
1. `try()`
2. `tryCatch()`
3. `withCallingHandlers()`

Again, `rlang` will provide some convenience functions for working with the condition system.

### Errors

We've seen errors before:

```{r}
f <- function() stop("This an error!")
f()
```
 If we don't want the call to be displayed:
 
```{r}
f <- function() stop("This an error!", call. = F)
f()

#in rlang we have the equivalent
f <- function() abort("Rlang error!")
f()
```

### Warnings

Warnings do not stop function evaluation like errors.

```{r}
options(warn = 0)
f <- function(){
  print("hi")
  warning("W1!")
  print("there")
  warning("W2!")
}

f()
```
```{r}
options(warn = 1) #print in place
f()
```
```{r}
options(warn = 2) #convert warnings to errors
f()
```

We have an `rlang` version of warnings:

```{r}
warn("uh oh") #same was warning with call. = F option
```


The `warnings()` command prints out the last set of warnings. If the are many warnings you can do `summary(warnings())`.

### Messages

Messages are strictly informational. 

```{r}
message("Awesome")
```

----------------------------------

# History

```{r}
# knitr::opts_chunk$set(echo = TRUE)
# knitr::opts_chunk$set(error = TRUE)
# library(rlang)
# x <- env(y = env(z = "hi", f = function() w <<- z))
# env_print(x)
# env_print(x$y)
# environment(x$y$f)
# environment(x$y$f) <- x$y
# environment(x$y$f)
# x$y
# env_bind(x, w = NULL)
# env_print(x)
# x$y$f()
# env_print(x)
# env_bind(x, w = "a")
# x$y$f()
# x$w
# parent.env(x$y)
# rm(w)
# library(rlang)
# x <- env(y = env(z = "hi", f = function() w <<- z))
# environment(x$y) <- x
# environment(x$y$f) <- x$y
# #we need to have w in x for super assignment to write w in x (otherwise it
# #it will be in the global environment)
# env_bind(x, w = NULL)
# x$y$f()
# env_parent(x$y)
# env_parent(x$y)
# parent.env(x$y) <- x
# env_parent(x$y)
# x
# library(rlang)
# x <- env(y = env(z = "hi", f = function() w <<- z))
# parent.env(x$y) <- x
# environment(x$y$f) <- x$y
# #we need to have w in x for super assignment to write w in x (otherwise it
# #it will be in the global environment)
# env_bind(x, w = NULL)
# x$y$f()
# rm(w)
# library(rlang)
# x <- env(y = env(z = "hi", f = function() w <<- z))
# parent.env(x$y) <- x
# environment(x$y$f) <- x$y
# #we need to have w in x for super assignment to write w in x (otherwise it
# #it will be in the global environment)
# env_bind(x, w = NULL)
# x$y$f()
# x$y$z
# x$w
# f <- function(x) g(x)
# g <- function(y) h(y)
# h <- function(z) log(z)
# f("a")
# f <- function(x) g(x)
# g <- function(y) h(y)
# h <- function(z) log(z)
# f("a")
# traceback()
# f <- function(x) g(x)
# g <- function(y) h(y)
# h <- function(z) log(z)
# f("a")
# traceback()
# f <- function(x) g(x)
# g <- function(y) h(y)
# h <- function(z) {lobstr::cst();log(z)}
# f <- function(x) g(x)
# g <- function(y) h(y)
# h <- function(z) log(z)
# f("a")
# traceback()
# f <- function(x) g(x)
# g <- function(y) h(y)
# h <- function(z) {lobstr::cst();log(z)}
# f("a")
# env_bind_lazy(y, wait = {Sys.sleep(1); 1})
# system.time(print(y$wait))
# system.time(print(y$wait))
# env_bind_lazy(x, wait = {Sys.sleep(1); 1})
# system.time(print(x$wait))
# system.time(print(x$wait))
# env_bind_active(x, now = function() runif(1))
# y$now
# env_bind_active(x, now = function() runif(1))
# x$now
# env_bind_active(x, now = function() runif(1))
# x$now
# x$now
# env_parents(last = empty_env())
# f <- function() stop("This an error!")
# f()
# f <- function() stop("This an error!", call. = F)
# f()
# f <- function() stop("This an error!", call. = F)
# f()
# #in rlang we have the equivalent
# f <- function() abort("Rlang error!")
# f()
# f <- function(){
# print("hi")
# warning("W1!")
# print("there")
# warning("W2!")
# }
# f()
# options(warn=0)
# f <- function(){
# print("hi")
# warning("W1!")
# print("there")
# warning("W2!")
# }
# f()
# options(warn = 1)
# f()
# options(warn = 0)
# f()
# option(warn = 0)
# f <- function(){
# print("hi")
# warning("W1!")
# print("there")
# warning("W2!")
# }
# f()
# options(warn = 0)
# f <- function(){
# print("hi")
# warning("W1!")
# print("there")
# warning("W2!")
# }
# f()
# f()
# options(warn = 1)
# f()
# f()
# options(warn = 1)
# f()
# options(warn = 2)
# f()
# warn("uh oh")
# options(warn=0)
# warn("uh oh")
# warning("uh oh", call. = F)
# warning("uh oh")
# f <- function() warning("uh oh")
# f()
# message("Awesome")
```

